\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates}

### 2022-12-30 submission/R1/manuscript.Rmd
update

We investigated the effect of 10 different classic illusions:

@@

Participants were first given a brief demographic survey, which collected information regarding their age, gender, country of birth, ethnicity and highest attained education level. This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block. Following this first series of blocks, two personality questionnaires were administered, the *IPIP6* [24 items, @sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the *PID-5* [25 items, @hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism). Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials). In total, each participant underwent 1340 trials of which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key  (left/right, or up/down depending on the illusion type). For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest. All trials were required to be completed within a single-session. The task was implemented using *jsPsych* [@de2015jspsych] and was hosted on *Pavlovia* [(**https://pavlovia.org/**)](https://pavlovia.org/){.uri}. The set of instructions for each illusion type is available in the experiment code.

@@

We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when implausibly fast). For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean). Additionally, due to a technical issue, no personality data was recorded for the first eight participants.

@@

Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age), and personality traits.

>>>>

We investigated the effect of 10 different classic illusions [**inseert figure**]:

@@

Participants were first given a brief demographic survey, which collected information regarding their age, gender, country of birth, ethnicity and highest attained education level. This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block. Following this first series of blocks, two personality questionnaires were administered, the *IPIP6* [24 items, @sibley2011] - measuring 6 "normal" personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the *PID-5* [25 items, @hopwood2012] - measuring 5 "pathological" personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism). Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials). In total, each participant underwent 1340 trials of which they had to respond "as fast as possible without making errors" (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key  (left/right, or up/down depending on the illusion type). For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest. All trials were required to be completed within a single-session (total experiment duration: ~55 minutes). The task was implemented using *jsPsych* [@de2015jspsych] and was hosted on *Pavlovia* [(**https://pavlovia.org/**)](https://pavlovia.org/){.uri}. The set of instructions for each illusion type is available in the experiment code.

@@

We excluded 6 participants upon inspection of the average error rate (when close to 50%, suggesting random answers), and reaction time distribution (when implausibly fast relative to the typical RT distribution). For the remaining participants, we discarded blocks with more than 50% of errors (2.16% of trials), possibly indicating that instructions were misunderstood (e.g., participants focused on the shorter line instead of the longer one), and 0.76% trials with extreme response times (\< 125 ms or \> 4 SD above mean RT). Additionally, due to a technical issue, no personality data was recorded for the first eight participants.

@@

Finally, for each of the individual illusion sensitivity scores (10 illusion-specific factors and the general Factor *i*), we tested the effect of contextual variables (screen size, screen refresh rate), demographic variables (sex, education, age), and personality traits. It should be noted that the screen size is a proxy of the true screen size, measured using the number of pixels.
